---
title: "Interactive Vaccine Data Visualizations"
author: "Jiani Wang"
output:
  flexdashboard::flex_dashboard:
    theme:
      bootswatch: flatly
      version: 3
    base_font:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(plotly)
library(DT)
library(ggplot2)
library(dplyr)
```

```{r, include=FALSE}
set.seed(42)
mock_vaccine <- data.frame(
  Vaccine = rep(c("H1N1", "H5N1", "H7N9"), each = 50),
  Reactogenicity = round(c(
    rnorm(50, mean = 5.0, sd = 2.0),
    rnorm(50, mean = 3.0, sd = 1.5),
    rnorm(50, mean = 4.0, sd = 1.8)
  ), 2),
  Adjuvant = sample(c("MF59", "AS03", "None"), 150, replace = TRUE),
  GMT_FoldChange = round(c(
    rnorm(50, mean = 4.0, sd = 1.2),
    rnorm(50, mean = 2.5, sd = 0.8),
    rnorm(50, mean = 3.2, sd = 1.0)
  ), 2),
  Seroprotection_Rate = round(c(
    runif(50, 70, 90),
    runif(50, 30, 50),
    runif(50, 50, 70)
  ), 1)
)

mock_vaccine$Reactogenicity <- pmax(pmin(mock_vaccine$Reactogenicity, 10), 0)
```

Column {data-width=400}
-----------------------------------------------------------------------
### Reactogenicity by Vaccine Type

```{r}
p1 <- ggplot(mock_vaccine, aes(
  x = Vaccine,
  y = Reactogenicity,
  color = Vaccine,
  text = paste("Vaccine Type:", Vaccine,
               "<br>Reaction Severity:", Reactogenicity))
) +
  geom_boxplot(alpha = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.7) +
  labs(
    title = "Reactogenicity by Vaccine Type",
    x = "Vaccine Type",
    y = "Reaction Severity (Scale: 0–10)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right", plot.title = element_text(size = 14))

ggplotly(p1, tooltip = "text")
```

> Key Takeaways  
> H1N1 shows the highest variability and overall severity in reported reactions.  
> H5N1 generally results in the mildest side effects.  
> These differences highlight how vaccine type and formulation may influence tolerability.

Column {data-width=500}
-----------------------------------------------------------------------
### Immunogenicity Summary Table

```{r}
table_data <- mock_vaccine %>%
  select(Vaccine, Adjuvant, GMT_FoldChange, Seroprotection_Rate, Reactogenicity)

colnames(table_data) <- c(
  "Vaccine Type", 
  "Adjuvant Used", 
  "Immune Response (GMT Fold-Change)", 
  "Seroprotection (%)", 
  "Reactogenicity (0–10)"
)

datatable(
  table_data,
  escape = FALSE,
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE,
    dom = 'tip',
    class = 'compact stripe hover row-border order-column nowrap',
    columnDefs = list(list(className = 'dt-center', targets = "_all")),
    searchHighlight = TRUE
  ),
  rownames = FALSE
)
```

> Key Takeaways  
> H1N1 produces the highest immune response, followed by H7N9 and then H5N1.  
> Adjuvants like MF59 and AS03 appear frequently in high-response cases.  
> This table helps explore how vaccine-adjuvant combinations affect immune strength and side effect severity.

Column {data-width=300}
-----------------------------------------------------------------------
### Dataset Description
<div style="font-size: 14px">
<ul>
<li>This dataset contains 150 simulated observations generated in R.</li>
<li>Variables include vaccine type, adjuvant, GMT fold-change, seroprotection rate (%), and reactogenicity (0–10).</li>
<li>The data mimics realistic patterns from immunogenicity and reactogenicity studies.</li>
<li>The original dataset from NIH-funded VTEU clinical trials is not public.</li>
</ul>
</div>

### Real-World Impact
<div style="font-size: 14px">
<ul>
<li>Helps visualize tradeoffs between vaccine safety and efficacy.</li>
<li>Useful for informing public health communication and formulation strategy.<li>
<li>Source code[View GitHub Repository](https://github.com/jianiw516/DATA555FinalProject)<li>
</ul>
</div>
